---
title: "Pre-class work 2"
author: "Jackie Goldman"
date: "September 16, 2018"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(warning=FALSE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(results="hold")
knitr::opts_chunk$set(cache=F)
knitr::opts_chunk$set(digits=3, scipen=3)
```
# pre-class


Make sure you commit this often with meaningfull messages. 

### Background

The exponential distribution is defined by its cumulative distribution function
\(F(x) = 1-e^{-\lambda x}\)

The R function ***rexp()*** generates random variables with an exponential distribution. For example 
<center><strong>rexp(n=10, rate=5)</strong> </center>

results in 10 exponentially distributed numbers with a rate \(\lambda=5\). If you leave out the 5 and just have
<center><strong>rexp(n=10) </strong></center>
then this results in 10 exponentially distributed numbers with a rate \(\lambda=1\), this is also referred to as the "standard exponential distribution". 

### Part 1


1. Generate 200 random values from the standard exponential distribution and store them in a vector `exp.draws.1`.  Find the mean and standard deviation of `exp.draws.1`.
```{r}
#200 observations, rate of 1
exp.draws.1<-rexp(200, rate=1)
mean(exp.draws.1)
sd(exp.draws.1)

```

2. Repeat, but change the rate to 0.2, 5, 7.3 and 10, storing the results in vectors called  `exp.draws.0.2`,  `exp.draws.5`,  `exp.draws.7.3` and  `exp.draws.10`. 
```{r}
#200 observations, rate of 0.2
exp.draws.0.2<-rexp(200, 0.2)
mean(exp.draws.0.2)
sd(exp.draws.0.2)

#200 observations, rate of 5
exp.draws.5<-rexp(200, 5)
mean(exp.draws.5)
sd(exp.draws.5)

#200 observations, rate of 7.3
exp.draws.7.3<-rexp(200, 7.3)
mean(exp.draws.7.3)
sd(exp.draws.7.3)

#200 observations, rate of 10
exp.draws.10<-rexp(200, 10)
mean(exp.draws.10)
sd(exp.draws.10)

```


3. The function `plot()` is the generic function in R for the visual display of data. `hist()` is a function that takes in and bins data as a side effect. To use this function, we must first specify what we'd like to plot.
    a. Use the `hist()` function to produce a histogram of your standard exponential distribution. 
    b. Use `plot()` with this vector to display the random values from your standard distribution in order.
    c. Now, use `plot()` with two arguments -- any two of your other stored random value vectors -- to create a scatterplot of the two vectors against each other.
```{r}
#question 3a
hist(exp.draws.1)

#querstion 3b
plot(exp.draws.1)
plotstandard
#question 3c
plot(exp.draws.0.2, exp.draws.7.3)

```

4. We'd now like to compare the properties of each of our vectors. Begin by creating a vector of the means of each of our five distributions in the order we created them and saving this to a variable name of your choice. Using this and other similar vectors, create the following scatterplots and explain in words what is going on:
    a. The five means versus the five rates used to generate the distribution.
    b. The standard deviations versus the rates.
    c. The means versus the standard deviations.

For each plot, explain in words what's going on.
### Question 4a
```{r}
means<-c(mean(exp.draws.1), mean(exp.draws.0.2), mean(exp.draws.5), mean(exp.draws.7.3), mean(exp.draws.10))
rates<-c(1, 0.2, 5, 7.3, 10)
plotmeans<-plot(means, rates )
```
In this first plot, we can see that the mean of the distribution decreases as the rate increases. The plot follows an exponential distribution.

### Question 4b

```{r}
StandardDevation<-c(sd(exp.draws.1), sd(exp.draws.0.2), sd(exp.draws.5), sd(exp.draws.7.3), sd(exp.draws.10))
plotsd<-plot(StandardDevation, rates)

```
The distribution of standard deviations plotted agains the rate looks similar to that of the plot of the means versus the rates. And similarly, as the rate increases, the standard devition decreases following an exponential distribution

### Question 4c
```{r}
meanvstandard<-plot(means, StandardDevation)
```
It looks as though there may be a linear relationship between the means and the standard deviations. For those distributions with small means, the standard deviation is also small. There is a clustereing around 0, and then points at 1 and 5. 
### Part II (PHP 2560 Only)


5. R's capacity for data and computation is large to what was available 10 years ago. 
    a. To show this, generate 1.1 million numbers from the standard exponential distribution and store them in a vector called`big.exp.draws.1`. Calculate the mean and standard deviation.
    b. Plot a histogram of `big.exp.draws.1`.  Does it match the function \(1-e^{-x}\)?  Should it? 
    c. Find the mean of all of the entries in `big.exp.draws.1` which are strictly greater than 1. You may need to first create a new vector to identify which elements satisfy this.
    d. Create a matrix, `big.exp.draws.1.mat`, containing the the values in 
`big.exp.draws.1`, with 1100 rows and 1000 columns. Use this matrix as the input to the `hist()` function and save the result to a variable of your choice. What happens to your data?
    e. Calculate the mean of the 371st column of `big.exp.draws.1.mat`.
    f. Now, find the means of all 1000 columns of `big.exp.draws.1.mat` simultaneously. Plot the histogram of column means.  Explain why its shape does not match the histogram in problem 5b).

### Question 5a
```{r}
big.exp.draws.1<-rexp(1100000, rate=1)
mean(big.exp.draws.1)
sd(big.exp.draws.1)

```
### Question 5b
```{r}
hist(big.exp.draws.1)
hist(exp.draws.1)
```
This histogram of the original distribution and that of the large distribution have the smae shape. However,  the tail of the large distribution is much longer. We would expect the distributions to have the same shape because they have the same rate. 

### Question 5c
```{r}
library(dplyr)
tbl1<-tbl_df(big.exp.draws.1)

big.exp.greater1<-filter(tbl1, value>1)

meanbig<-mean(big.exp.greater1$value)
meanbig

```
  
### Question 5d
```{r}
big.exp.draws.1.mat<-matrix(big.exp.draws.1, ncol=1000, nrow=1100)
bighist<-hist(big.exp.draws.1.mat)
```
###Question 5e
```{r}
c371<-big.exp.draws.1.mat[,371]
mean(c371)
```
#Question 5f
```{r}
columnMeans<-colMeans(big.exp.draws.1.mat[], dims=1)
hist(columnMeans)
```
The histogram is normally distributed. Though the distribution that generated this final matrix follows an exponential distribution, after creating a matrix with 1,000 columns, comprising of 1,100 rows, the column means become normally distributed. 

