# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}
for (i in 1:50) {
  print(i)
  i = i+1
}
```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
#initiatize sum to 0;
sum<-0;
#now add
for (i in 1:1000) {
  sum<-sum + i
}
sum
```

B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
seq<-seq(2, 1000,2)
sumb<-0
for (i in seq){
  sumb<-sumb + i
}
sumb

```

C. Now, repeat A and B WITHOUT using a loop.
```{r}
#Question 2A
all <- seq(from=0, to=1000, by=1)
#summing vector
sum(all)
#or using arthmatic
1000*(1000+1)/2

#Question 2B
#making vector with even integers
even <- seq(from=0, to=1000, by=2)
#summing vector
sum(even)
#now using arithmatic
1000*(1000+2)/4
```
### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.
```{r}
#creating datafram
survey.clean <- survey

#cleaning dataframe
for (i in 1:6) {
  for (j in 2:6 ){
  if (survey[i,j] >1 & survey[i,j] <5) {
    survey.clean[i,j] = survey[i,j]
  } else {
    survey.clean[i,j] = NA
}
  }
}
```
B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
#creating new column
survey.clean$invalid.answers <- rep(0)

#adding number of invalid answers
for (i in 1:6) {
  for (j in 2:6 ) {
    if (is.na(survey.clean[i,j]) == TRUE) {
      survey.clean[i,7] = survey.clean[i,7] + 1
    }
    }
  }
```
