Pre-Class Problems
Use dplyr functions to address the following questions:
### Question 1
How many unique countries are represented per continent?
```{r}
install.packages("gapminder")
library(dplyr)
library(gapminder)
gapminder
tbl_gpm<-tbl_df(gapminder)
gapminder %>%
  group_by(continent)%>%
  summarise(n_distinct(country))
```
There are 52 countries in Africa, 25 listed in the Americas, 33 in Asiam 30 in Europe and 2 in Oceania listed in the dataset.

### Question 2
Which European nation had the lowest GDP per capita in 1997?
```{r}
gapminder %>%
  filter(continent=="Europe" & year=="1997") %>%
  arrange(gdpPercap) 
```
The country in Europe in 1997 with the lowest GDP per capita is Albania

### Question 3
According to the data available, what was the average life expectancy across each continent in the 1980s?
```{r}
gapminder %>%
  filter(year>=1980 & year<=1989) %>%
  group_by(continent) %>%
  summarise(mean_life_exp=mean(lifeExp, na.rm=TRUE))

```
The mean life expectancy throughout the 1980's is 52.5 years in Africa, 67.16 years in the Americas, 63.73 years in Asia, 73.22 years in Europe, and 74.81 years in Oceania.
### Question 4
What 5 countries have the highest total GDP over all years combined?
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(total_gpb=sum(gdpPercap, na.rm=TRUE)) %>%
  arrange(desc(total_gpb))

```
The 5 countries with the highest GDP per capital are Kuwait, Switzerland, Norway, the United States, and Canada.

### Question 5
What countries and years had life expectancies of at least 80 years? N.b. only output the columns of interest: country, life expectancy and year (in that order).
``` {r}
gapminder %>%
  filter(lifeExp>=80) %>%
  select (lifeExp, country, year)
```
### Question 6
What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
gapminder %>%
  group_by(country) %>%
summarize(cor=cor(lifeExp, gdpPercap)) %>%
  arrange(desc(abs(cor)))

```
The countries with the 10 highest correlations between life expectancy and gdp per capita are France, Austria, Belgium, Norwar, Oman, United Kingdom, Italy, Israel, Denmark and Australia.
### Question 7
Which combinations of continent (besides Asia) and year have the highest average population across all countries? N.b. your output should include all results sorted by highest average population. With what you already know, this one may stump you. See this Q&A for how to ungroup before arrangeing. This also behaves differently in more recent versions of dplyr.
```{r}
gapminder %>%
  group_by(continent, year) %>%
  mutate(rank=row_number(pop)) %>%
  filter(rank<=10) %>%
  arrange(rank)%>%
  ungroup() %>%
  arrange(desc(pop))
```
### Question 8
Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data?
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(sd=sd(pop)) %>%
  arrange(sd)

```
The populations with the most consistent populations are Soa Tome and Principe, Iceland and Montenengro. 
###Question 9
Subset gm to only include observations from 1992 and store the results as gm1992. What kind of object is this?

```{r}
gm1992<-gapminder %>%
  filter(year==1992)

```
gm1992 is a data frame.

### Question 10
Which observations indicate that the population of a country has decreased from the previous year and the life expectancy has increased from the previous year? See the vignette on window functions.
```{r}
gapminder %>%
  mutate(pop_delta=pop-lag(pop)) %>%
  mutate(lifeExp_delt=lifeExp-lag(lifeExp)) %>%
  filter(pop_delta <0, lifeExp_delt>0)
           
```
This code produces a table where the countries have a positive increase in life expectancy from the previous year and a decreased population from the year before. 