# pre-class


Make sure you commit this often with meaningful messages. 



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.

``` {r}
f1 <- function(string, prefix) { #this function operates on a sting and tests whther each element of the character vector starts with the prefix listed in the code. It would retrn a true false. I would name it prefix_true_false
  substr(string, 1, nchar(prefix)) == prefix #this line of code extracts the string, and truncates the string to 
}
f2 <- function(x) { #this function takes x, if the length of x is less than or equal to one, it retunrs a NULL anser, and if it is greater o equal to one, it removes the last element. I might name it remove_last
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f3 <- function(x, y) { #this function repeates y once for each elemenet of x, I might call it duplicate_second
  rep(y, length.out = length(x))
}
```

2. Compare and contrast rnorm() and MASS::mvrnorm(). How could you make them more consistent?


3. Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset. 

4. Use vapply() to:
    a. Compute the standard deviation of every column in a numeric data frame.
    b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
